Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections

### 2025-07-07 response_memo.qmd

---
title: "Response Memo"
format: 
  hikmah-response-typst:
    toc: false
    mainfont: TeX Gyre Pagella
    fontsize: 11pt
    bibliography: references.bib
    bibliographystyle: american-political-science-association.csl
---

I thank the Reviewers for taking time to provide their thoughtful and constructive comments. I revised much of the initial manuscript by restructuring certain sections and added clarification where needed. The Reviewers' comments are cited or summarized in [red italicized text]{.memo-reviewer-inline}, and my responses are in black text. I include excerpts from my revised manuscript in [blue]{.memo-excerpt-inline}. Note that an ellipses ($...$) within an excerpt of the manuscript stands in for the main text.

@@

# General Reviewer concerns on theory and competing explanations

Both reviewers' concerns touched on the same general issue, and the offered critiques and recommendations complemented each other. [*Reviewer 1*]{.memo-reviewer-inline} notes a general lack of discussion on potential explanations or explicit conjecture on a potential mechanism. [*Reviewer 2*]{.memo-reviewer-inline} recognized the same as ambiguity within of the theory; synthesis of the context (i.e., background section) and the literature didn't lead to a clear and concise reason for "why veterans" would be expected to influence confidence in elections. I dedicated a separate section titled as such hoping to clarify the source of my stated expectations, however I agree that the issue needs to be expounded at both ends.

These additions and revisions are extensive, yet I believe adding a refined theory section and discussion section addresses the most immediate concerns raised by both reviewers. The theory section is revised to make clear "why veterans", noting the gaps in available literature while making the intent of the current study explicit. The added discussion section complements the theory section by discussing potential explanations.

## ***Reviewer 1***---***what is the mechanism here?***

::: memo-reviewer
We understand that veterans are generally viewed positively, and that’s part of why this approach might work. However, in the data, we don’t see any direct measures of how respondents feel about veterans—there’s no pre-treatment baseline for that. I understand that the data might not allow for it, or that asking upfront could interfere with the treatment, but still---it leaves us wondering: what exactly is driving the results? Is it the idea of integrity, a feeling of safety or strength, or a sense of national service? We don’t really know. For this paper, it might be acceptable to say we’re building on prior work and that the cue is sufficient, but it would be helpful to acknowledge that gap more directly and discuss potential explanations.
:::

@@

In response to [*Reviewer 1*]{.memo-reviewer-inline}'s comment above, I included the following text in the '**Limitations**' section,

::: memo-excerpt
## Limitations

The bluntness of the experimental stimulus and survey items don't allow us to establish exactly what it is about military veterans in particular that cues stronger confidence in elections. No survey items assessed public attitudes about military veterans specifically, pre- nor post-treatment. This unfortunately leaves no room for me to postulate and examine any potential mechanisms underlying results.
:::

Also, I extended a discussion prior to the conclusion on potential explanations and avenues for future research.

::: memo-excerpt
## Discussion

```{=typst}
#sym.dots.h.c
```
A question lingers about whether some antecedent mechanism transmits (i.e., mediates) the effects of the veteran treatment vignette to the outcome. To put simply, this study reiterates that there *is* something special about veterans in particular that elicits greater confidence in elections, but cannot go further in determining exactly *what* that special something might be. Factors that *moderate* the relationship of the treatment on the outcome of interest are likely plenty, but one can only offer conjecture of a *mediator* driving results.

@@

One potential explanation is that veterans, and all those who serve in the military, potentially function as a national symbol. The group is abstract just as the nation is abstract, and the potency of the veteran group emulates the potency of a national symbol. Recall Benedict Anderson's [-@anderson2006] meditation on the Tomb of the Unknown Soldier, "...saturated with ghostly *national* imaginings" [-@anderson2006, 9, emphasis original]. In various studies, national symbols such as the flag, the national anthem, or even mere appeal to national identity have potent, yet variable, effects on different outcomes [@schatz2007; @butz2007; @kemmelmeier2008; @butz2009; @gangl2016; @kalmoe2016; @levendusky2018]. It is plausible that military service members and veterans function in much the same way---as a symbol of national loyalty, pride, and nationalism for people to praise and revere out of a sense of duty prescribed by social norm and reinforced by national sentiment. As such, it may be the case that the treatment effect on confidence in elections is mediated through an affective response to a national symbol---in this case, veterans. The veteran cue may indeed stimulate national sentiment and subsequently boost confidence in elections.

In results presented here, confidence in elections is associated with greater confidence among those in the treatment group, but neither vignette conveys information about the procedures or practices meant to safeguard election integrity; nothing in the treatment suggests to respondents that election integrity would be improved by recruiting veterans. Arguably, the results presented here speak to the "emotional pathway" mentioned in @gaudette2025 so long as we assume that there's an inherent appeal to national sentiment (e.g., nationalism, patriotism) by use of veterans[^1]. Importantly, if the veterans treatment is mediated by stimulation of affect for nation or national identity, then comparable stimulation of national sentiment unrelated to veterans should produce similar results via this "emotional-pathway". However, appeals to national sentiment by use of sterile national symbols (e.g., banal presentation of the U.S. flag, playing of the national anthem) may not have the same affective influence on confidence in elections.

@@

In a similar vein, it is plausible that certain political apprehensions are nullified insofar as veterans as a group are generally perceived as 'apolitical'---which itself may be enough to engender favor among the mass public [@krupnikov2022]. Current and former military service members may be viewed by the public generally as 'apolitical', especially in comparison to practically any other social group. Indeed, an 'apolitical' reputation has been long desired by military leaders and top brass [@lagrone2019; @garamone2016; @garamone2016a; @baldor2020] despite evidence to the contrary [@betros2001; @mcnerney2006]. Nevertheless, the putative 'apolitical' nature of military service may generalize to veteran service members, and increase confidence in elections based on the assumption that veterans uphold a commitment to ensure election integrity above their own personal loyalties, partisanship, or politics. Prior research has relied on the assumption that such a "*group-based*" norm persists among military members, which prescribes that the military is, or ought to be, apolitical [@mullinix2022]. Although, the extent to which this perception generalizes widely among the mass public about the military in general, or about current and former service members in particular, has yet to be examined. Given recent events of far more overt politicization of the U.S. military [@baldor2025; @baldor2025a; @baldor2025b; @baldor2025c; @thomas2025], it would be useful for future research to consider whether perceptions about the military generally, and military leaders in particular, generalize to perceptions about current and former military service members. Especially insofar as such attitudes relate to confidence in the electoral process and administration.
:::

## ***Reviewer 1***---***Hypothesis structure could be clearer***

::: memo-reviewer
You present H1 very clearly, which is great. However, later in the paper, there are additional expectations that seem like they could have been articulated as formal hypotheses too—perhaps H2, H3, or even sub-parts like H1a, H1b. While it’s not a major issue, clearer hypothesis structure would enhance readability and help clarify the paper’s logic as it progresses. Currently, it feels like a mix of formal and informal framing.
:::

@@

Agreed. I revised and inserted the following hypotheses more formally, and also inserted text to make more explicit the reasoning behind the latter hypotheses. The text of the '**Hypotheses**' section is unchanged save for inclusion of the following text.

::: memo-excerpt
## Hypotheses

```{=typst}
#sym.dots.h.c
```
> H~2~: Differences in confidence in elections in one's local area and confidence in elections in Maricopa County, AZ, will be smaller among those who are presented with an announcement that election officials are recruiting miilitary service veterans to work as election staff and volunteers compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

```{=typst}
#sym.dots.h.c
```
> H~3~: Among those who refute the legitimacy of the 2020 election results, those presented with an announcement that election officials are recruiting veteran service members for election jobs will be associated with greater levels of confidence in elections compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

@@

Prior research has found that political candidates benefit from having a military background [@mcdermott2015], especially evaluations of candidate's prospective performance on particular war issues (e.g., foreign policy, terrorism) [@hardy2019]. The idea is that prior military service provides meaningful information to use as a heuristic in regard to such "war issues". Although veterans are no longer actively serving, there's potentially some lingering general assumptions about the capacities and motives of military veterans to provide safety and security arising from the nature of their prior occupation as members of the U.S. armed forces. Insofar as the physical safety of people at election sites is a "military matter", then the potential that election sites may be staffed with military veterans may ease related concerns about the safety and security of those places. As such, I expect that concerns about potential for violence will be lower, and confidence in voter safety higher, when presented with information that veterans are being actively recruited to work and volunteer in election offices.

> H~4a~: Concerns about the potential for violence, threats of violence, or intimidation while voting in person will be lower among those presented with an announcement that election officials are recruiting veteran service members for election jobs compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

> H~4b~: Confidence that voters will be safe to vote in-person at election sites will be higher among those presented with an announcement that election officials are recruiting veteran service members for election jobs compared to those presented with an announcement of recruitment efforts that do not mention military veterans.
:::

## ***Reviewer 1***---***Racial dynamics and trust***

::: memo-reviewer
This aspect also emerged for me. The sample is mostly white, making it challenging to draw conclusions about how this might work for communities of color. As we know, these communities often have long histories of institutional distrust—including towards the military. Thus, it’s possible the cue might not resonate in the same way, or perhaps it does, as many veterans are people of color too. We don’t know, and the current data cannot address that limitation—but acknowledging it and speculating could be beneficial.
:::

@@

I included the following text in the discussion to address [**Reviewer 1**]{.memo-reviewer-inline}'s comment above.

::: memo-excerpt
The data of the current study is limited in helping to discern whether the treatment garners greater confidence in elections uniformly across different racial or ethnic groups given that the largest proportion of the sample identified their race as White (77%). It is reasonable to suspect that the treatment vignette might differ by race and ethnicity, but the sample is insufficient to make such generalizations. Nevertheless, attitudes about veterans that differ by group may play an important role in determining whether publicized efforts to recruit military veterans boosts confidence in elections uniformly or whether the effect is merely partial to certain segments of the population.

It is definitely plausible that the public, to some unknown extent, associates a particular gender or race, or both, with former and active military members which may, in turn, mediate the effect of the veteran cue. Additionally, I can't be sure as to whether respondent's views of the military generalize to attitudes about active service members and veterans. However that is a separate question about the source of public opinion about military members themselves beyond the scope of this study.

There's scant research or evidence available to postulate a well-formed theory on the mechanism underlying the expected effects of veterans as a group on attitudes about institutions and processes, rather than on attitudes about persons or groups. That is to say, there's not much research on how the public feels, or what they think, about former members of the U.S. Armed Forces.

In their research examining whether groups differed in attitudes about veterans, @kleykamp2018 could only speculate as to why levels of support were lowest among Hispanic and Black respondents. Their study was distinctive in that it stands as a rare case in which public attitudes about veteran military members was the specific focus of inquiry as opposed to studies that examine public opinion about the military in general. Other research has found that attitudes about the U.S. military generally---though not about active or veteran military service members in particular---differs by race and ethnicity to varying extents, but also could not supply reason for one groups' sentiment over another^[Note, however, that the sample examined by @nichols2015 was constrained to Texas residents, and the military-related issues concerned, among other things, favor for reinstatement of a military draft, support for allowing undocumented immigrants who live in the U.S. to serve in the U.S. Armed Forces, and whether undocumented immigrants should be granted U.S. citizenship as a result of military service. Although the authors found high overall support for the military as an institution, the other military-related attitudes of interest diverge far from public opinion about veterans. As such, differential effects found among different racial or ethnic groups likely have little to do with attitudes about former or active military service members.] [@nichols2015].

Other studies have drawn out disparities in public attitudes about veterans but in ways that are mostly indirect, i.e., without assessing public regard for veterans directly. One overarching issue is that there's limited research that differentiates public regard for veteran and active military service members from the U.S. military as an institution in general. Some studies assess public regard for veterans indirectly by examining whether, and to what extent, endorsement of certain policies differentiate by whether such policies benefit veterans or the public generally. In one such study by @tsai2021, Republican partisanship was shown to be the most consistent factor associated with more favorable attitudes (e.g., compassion) toward homelessness and PTSD among the veteran population, but not among the general adult population. Additionally, Republican partisans tended to show greater support for federal funding for homelessness and PTSD research when the reference population was military veterans. That study, however, suggests that the effect of the veteran status cue is sensitive to the topic in question.

Perhaps unsurprisingly, military veterans appear to be uniquely deserving in the eyes of Republican partisans [@tsai2021; @mcdermott2015; @richardson2022]. As such, military veteran status may qualify favorable attitudes among Republican partisans whereas their regard for other non-veteran groups might otherwise express contempt or indifference. This further indicates a need for research to determine an underlying mechanism explaining why, and for whom, veterans incur special favor.
:::

## ***Reviewer 1***---***How generalizable are these results***

::: memo-reviewer
I appreciate your honesty about how the strongest effects appear in Maricopa County, and that we don’t observe the same impact when people consider their own local elections. That’s interesting in itself, but it raises questions—can this approach be effective in other areas? Would appointing veterans as officials in lower-salience places have any impact? Or is this cue only meaningful in scrutinized locations? You touch on this briefly, but expanding this section would add depth, particularly for readers contemplating its implications for future elections.
:::

@@

To address these concerns, I include the following text in the '**Limitations**' section

::: memo-excerpt
The setting of the vignette and specified in survey items might raise additional concerns on whether results are generalizable. Naming a specific county in the vignette adds in a potentially influential factor unaccounted for in the survey; some people may have attitudes about the county in question, others won't, while others may be muffed to consider a random county in the U.S. they know nothing about. Naming a real and somewhat salient county potentially undermines confidence that the treatment effects are solely attributable to veterans to an unknown degree. Simply put, there may be something special about Maricopa County, AZ in particular that generalizes across attitudes of trust or distrust in elections.

@@

As stated previously, the survey was fielded on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. The results presented here are not assumed to be generalizable to the national population. It may be feasible to weigh the sample to demographics of the population at the time the survey was fielded, but this is unnecessary for the experimental purpose of the study. If similar findings complement or replicate these results, then the expense of drawing samples more representative of the national population may be justified. At this stage, however, the results are limited to demonstrating treatment effects.

Note also that the measurement items of the components trust and distrust, and the method of computing scores for the dependent variable, confidence in elections, are unique to this study but falls under the methodological framework of Classical Test Theory (CTT). Although survey measurement items were either directly drawn from or inspired by similar survey items found in various other surveys (e.g., SPAE, Pew), the question wording, response choices, coding, and computation of sum scores of the dependent variable are specific, thus rendering measurement of the hypothetical constructs of interest (i.e., trust, distrust) less comparable across similar studies that profess measurement of similar constructs [@widaman2024]. Namely, trust, distrust, or confidence in elections as defined and measured here.
:::

## ***Reviewer 2***---***Theory***

::: memo-reviewer
The author could be clearer about the puzzle: Why would identifying poll workers as veterans plausibly increase confidence in elections? Why might it not? The paper gestures at answers---trust, apolitical reputation, shared identity---but doesn’t quite fully develop the competing expectations.
:::

In response to [*Reviewer 2*]{.memo-reviewer-inline}'s comments concerning the theory, I restructured the theory section ("**Why veterans**"), expand on the premises, and add literature that better leads to the expectations. The text below reflects the revised theory section that better incorporates "Why Veterans".

::: memo-excerpt
## Theory: Why Veterans

The theory tested in this study is this: when the public is informed that election officials are actively recruiting veteran military service members to work at election sites as staff or volunteers, their overall confidence in the fairness, accuracy, and integrity of the electoral process will improve. Said another way, insecurities over the integrity of elections administration and safety of voters lessen as the public learns that military service veterans are being sought out by election officials to staff or volunteer at election sites. The stimulating influence of veterans as a group in the abstract is expected to counteract distrust and insecurity pertaining to the institution of elections administration, as well as inspire greater trust in the same. This prompts an obvious question: why would information about efforts to recruit veterans to work as election staff and volunteers have any influence on confidence in elections?

Election officials don't focus on recruiting people of a particular demographic, status, class, or prestige in order to shore up public trust in elections, and they don't discriminate recruitment efforts based on who the public thinks is best suited to secure elections from fraud. Election officials are likely to be agnostic as to who dedicates their time to civil service such as election work. Especially since staffing has been an issue since at least the 1990s [@maidenberg1996; @ferrer2024]. In 2020, such efforts were made far more difficult by the COVID-19 pandemic [@mena2020; @abbate2020a]. There's no special reason to target veterans for recruitment above other groups. Indeed, there's no reason to discriminate recruitment efforts at all if the point is purely to fill staffing vacancies.

There was a sudden push to target veterans for recruitment efforts $...$

One can speculate that $...$

The claim that recruiting veterans to work as election staff and volunteers would improve public confidence in elections emerges as non-profit organizations---mainly comprised of veteran and military family members themselves---operate under, and promote, that assumption [@looker2024; @lawrence2024; @wentling2024]. Non-profit groups of late have made incredible strides in targeted recruitment efforts under the notion that veterans are uniquely capable, committed, and best fit to serve the needs of the United States as civil servants and volunteers at election sites. Indeed, this study can be seen as a test of a particular assumption that veterans have about themselves---the assumption that their mere presence and civic engagement will inspire confidence in elections among the public.

Veterans are, after all, civilians. Civilians with shared (but varied) experience and status of prior military service, but who are no more or less qualified for this civil service[^3] over non-veteran civilians. To be frank, announcements about concerted efforts to recruit veterans as election workers should not reasonably influence one's trust that votes will be counted more accurately, or that election fraud will not occur; there is no known evidence demonstrating that actually employing veterans as election staff or volunteers further improves the safety, security, and overall integrity of elections in the United States. Moreover, there's no known evidence of positive influence on public perceptions of said outcomes when voters encounter military veterans working as staff or volunteers at elections sites. So, why veterans?

Two considerations underpin the notion that information about veterans will improve trust and quell distrust in elections. First, positive public regard for veterans *in general* is sufficient to elicit greater trust and lower distrust in elections despite the fact that information about veterans as election workers, or even merely information on efforts to recruit veterans, has no bearing on the actual integrity of the electoral process.

### 2025-07-07 docs/_tex/index.tex
revisions

Regardless of the measures taken by election officials to boost public
confidence in the \emph{trustworthiness} of election administration
(e.g., conducting audits, testing election machines), public
\emph{trust} and confidence in elections more generally is apt to shift
dramatically post-election based on factors such as partisanship, elite
rhetoric, particular state policies, and more
(\citeproc{ref-carter2024}{Carter et al. 2024};
\citeproc{ref-coll2024a}{Coll and Clark 2024};
\citeproc{ref-nadeau1993}{Nadeau and Blais 1993}). Moreover, prior
research has found that evaluation of election workers themselves are an
important factor when it comes to levels of public confidence in the
electoral process (\citeproc{ref-claassen2008}{Claassen et al. 2008};
\citeproc{ref-hall2007}{T. Hall, Monson, and Patterson 2007};
\citeproc{ref-hall2009}{T. E. Hall, Quin Monson, and Patterson 2009}).
Such studies focused on the quality of the voter experience with
reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality
of the voter experience may be influenced merely by prior impressions
about \emph{who} comprises election staff and volunteers. Political and
other social science researchers have recognized for some time the power
that group identity can have over attitudes and perception
(\citeproc{ref-vanbavel2021}{Van Bavel and Packer 2021};
\citeproc{ref-xiao2016}{Xiao, Coppin, and Van Bavel 2016};
\citeproc{ref-xiao2012}{Xiao and Van Bavel 2012}). As such, we can
expect that information identifying the particular groups being
recruited to serve as election staff will be enough to improve
confidence in elections, and lessen expectations of electoral fraud. The
next section elaborates on the proposed theory, terms utilized, and why
military veterans are of particular interest in this regard.

\section{Theory}\label{theory}

This theory is constrained to public confidence in the administration of
elections in the United States. To be clear, the scope conditions in
which this theory applies concern public trust and distrust in election
administration in the anticipatory period prior to the election event.
Moreover, this study is, to a limited extent, indirectly related to
public regard about veteran members of the U.S. Armed Forces.

@@

The theoretical relationship between trust and distrust is important to
understand, especially when it comes to the dependent outcome of
interest. Before going further, it is necessary to explicate the
conception \emph{confidence in elections} with respect to trust and
distrust.

Trust is, after all, merely \emph{belief} held in the face of
uncertainty. Trust is reliance on the veracity of belief in the face of
uncertainty. Said another way, trust is assured belief dependent on its
supposed veracity; a state of certainty with respect to some belief or
judgement. It is ``assured'' because it is one's belief that relies on
something rather than nothing. Hence quantitative methods of the
sciences are replete with things such as \emph{confidence} intervals
around our estimates. Therefore, I understand trust to refer to the
state or quality of certainty around one's beliefs or expectations; the
extent to which one is assured of their expectations regarding what is
or what will be the case.

@@

\subsection{Why Veterans}\label{why-veterans}

Election officials don't focus on recruiting people of a particular
demographic, status, class, or prestige in order to shore up public
trust in elections, and they don't discriminate recruitment efforts
based on who the public thinks is best suited to secure elections from
fraud.

Election officials are likely to be agnostic as to who dedicates their
time to civil service such as election work. Especially since staffing
has been an issue since at least the 1990s
(\citeproc{ref-ferrer2024}{Ferrer, Thompson, and Orey 2024};
\citeproc{ref-maidenberg1996}{Maidenberg 1996}). In 2020, such efforts
were made far more difficult by the COVID-19 pandemic
(\citeproc{ref-abbate2020a}{Abbate 2020}; \citeproc{ref-mena2020}{Mena
2020}). There's no special reason to target veterans for recruitment
above other groups. Indeed, there's no reason to discriminate
recruitment efforts at all if the point is purely to fill staffing
vacancies.

Veteran military service members are, after all, civilians; Civilians
with shared (but varied) experience and status of prior military
service, but who are no more or less qualified for this civil
service\footnote{Research is limited and mixed on whether veterans in
  general are more inclined, however (\citeproc{ref-nesbit2011}{Nesbit
  and Reingold 2011}).} than non-veteran civilians. To be frank,
announcements about concerted efforts to recruit veterans as election
workers should not reasonably influence one's trust that votes will be
counted more accurately, or that election fraud will not occur; there is
no known evidence demonstrating that actually employing veterans as
election staff or volunteers further improves the safety, security, and
overall integrity of elections in the United States. Moreover, there's
no known evidence of positive influence on public perceptions of said
outcomes when voters encounter military veterans working as staff or
volunteers at elections sites.

Rather, the claim that recruiting veterans to work as election staff and
volunteers would improve public confidence in elections emerges as
non-profit organizations---mainly comprised of veteran and military
family members themselves---operate under, and promote, that assumption
(\citeproc{ref-lawrence2024}{Lawrence 2024};
\citeproc{ref-looker2024}{Looker 2024};
\citeproc{ref-wentling2024}{Wentling 2024}). Non-profit groups of late
have made incredible strides in targeted recruitment efforts under the
notion that veterans are uniquely capable, committed, and best fit to
serve the needs of the United States as civil servants and volunteers at
election sites. Indeed, this study can be seen as a test of a particular
assumption that veterans have about themselves---the assumption that
their mere presence and civic engagement will inspire confidence in
elections among the public.

The key findings of literature on confidence in elections hardly point
to public regard for veterans as a mechanism for the theoretical
expectations of improving overall confidence in elections and improving
feelings of voter safety at election sites. Comparable research finds
that confidence in elections is increased through elite rhetoric from
fellow partisans affirming legitimacy of elections
(\citeproc{ref-clayton2023}{Clayton and Willer 2023}), or through the
provision of information about the electoral process or other official
messaging of election officials themselves
(\citeproc{ref-gaudette2025}{Gaudette et al. 2025}). Neither of the
studies just cited, however, consider attitudes or sentiment about
veterans as a source for dispelling distrust and improving trust in
elections. To be clear, there is no immediate ``puzzle'' in the
literature that immediately draws out public sentiment for military
veterans as a source for bolstering confidence in elections. That being
said, the general public perception, attitudes, or even stereotypes
about military veterans are significant to consider.

It is important to make clear two considerations that underpin the
notion that information about veterans will improve trust and quell
distrust in elections. First, positive public regard for veterans
\emph{in general} is sufficient to elicit greater trust and lower
distrust in elections despite the fact that information about veterans
as election workers, or even merely information on efforts to recruit
veterans, has no bearing on the actual integrity of the electoral
process. An underlying assumption here is that there \emph{is} something
special about veterans, but this doesn't include a proposition on
exactly \emph{what} that special something might be in theory. This is
partly due to the limited research available on public regard for
veterans, specifically in terms of elucidating an underlying mechanism
for such positive favor. Moreover, there is limited research on public
opinion about former and active military service members distinguished
separately from public opinion about the military as an institution in
general. Endorsement or support of government policies intended to
benefit veterans specifically (e.g., federal funding on veteran
healthcare) are easy to conflate as interchangeable with public
attitudes about veterans as a group in particular. This, in turn,
permits general expectations of positive influence due to the public's
overarching positive regard for veterans, but disallows for expectations
regarding how respondents would actually process and interpret a
treatment that only mentions efforts to recruit them for election work.
In short, we are sure that people sustain positive views for veterans
generally, but beyond that I make no special effort to propose and
examine why that is the case in greater detail.

>>>>

Regardless of the measures taken by election officials to boost public
confidence in the \emph{trustworthiness} of election administration
(e.g., conducting audits, testing election machines), public
\emph{trust} and confidence in elections more generally is apt to shift
dramatically post-election based on factors such as partisanship, elite
rhetoric, particular state policies, and more
(\citeproc{ref-carter2024}{Carter et al. 2024};
\citeproc{ref-coll2024a}{Coll and Clark 2024};
\citeproc{ref-nadeau1993}{Nadeau and Blais 1993}). Comparable research
finds that confidence in elections is increased through elite rhetoric
from fellow partisans affirming legitimacy of elections
(\citeproc{ref-clayton2023}{Clayton and Willer 2023}), or through the
provision of information about the electoral process or other official
messaging of election officials themselves
(\citeproc{ref-gaudette2025}{Gaudette et al. 2025}). Moreover, prior
research has found that evaluation of election workers themselves are an
important factor when it comes to levels of public confidence in the
electoral process (\citeproc{ref-claassen2008}{Claassen et al. 2008};
\citeproc{ref-hall2007}{T. Hall, Monson, and Patterson 2007};
\citeproc{ref-hall2009}{T. E. Hall, Quin Monson, and Patterson 2009}).
Such studies focused on the quality of the voter experience with
reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality
of the voter experience may be influenced merely by prior impressions
about \emph{who} comprises election staff and volunteers. Political and
other social science researchers have recognized for some time the power
that group identity can have over attitudes and perception
(\citeproc{ref-vanbavel2021}{Van Bavel and Packer 2021};
\citeproc{ref-xiao2016}{Xiao, Coppin, and Van Bavel 2016};
\citeproc{ref-xiao2012}{Xiao and Van Bavel 2012}). As such, we can
expect that information identifying the particular groups being
recruited to serve as election staff will be enough to improve
confidence in elections, and lessen expectations of electoral fraud. The
next section elaborates on the proposed theory and why military veterans
are of particular interest in this regard.

\subsection{Theory: Why Veterans}\label{theory-why-veterans}

The theory tested in this study is this: when the public is informed
that election officials are actively recruiting veteran military service
members to work at election sites as staff or volunteers, their overall
confidence in the fairness, accuracy, and integrity of the electoral
process will improve. Said another way, insecurities over the integrity
of elections administration and safety of voters lessen as the public
learns that military service veterans are being sought out by election
officials to staff or volunteer at election sites. The stimulating
influence of veterans as a group in the abstract is expected to
counteract distrust and insecurity pertaining to the institution of
elections administration, as well as inspire greater trust in the same.
This prompts an obvious question: why would information about efforts to
recruit veterans to work as election staff and volunteers have any
influence on confidence in elections?

Election officials don't focus on recruiting people of a particular
demographic, status, class, or prestige in order to shore up public
trust in elections, and they don't discriminate recruitment efforts
based on who the public thinks is best suited to secure elections from
fraud. Election officials are likely to be agnostic as to who dedicates
their time to civil service such as election work. Especially since
staffing has been an issue since at least the 1990s
(\citeproc{ref-ferrer2024}{Ferrer, Thompson, and Orey 2024};
\citeproc{ref-maidenberg1996}{Maidenberg 1996}). In 2020, such efforts
were made far more difficult by the COVID-19 pandemic
(\citeproc{ref-abbate2020a}{Abbate 2020}; \citeproc{ref-mena2020}{Mena
2020}). There's no special reason to target veterans for recruitment
above other groups. Indeed, there's no reason to discriminate
recruitment efforts at all if the point is purely to fill staffing
vacancies.

The claim that recruiting veterans to work as election staff and
volunteers would improve public confidence in elections emerges as
non-profit organizations---mainly comprised of veteran and military
family members themselves---operate under, and promote, that assumption
(\citeproc{ref-lawrence2024}{Lawrence 2024};
\citeproc{ref-looker2024}{Looker 2024};
\citeproc{ref-wentling2024}{Wentling 2024}). Non-profit groups of late
have made incredible strides in targeted recruitment efforts under the
notion that veterans are uniquely capable, committed, and best fit to
serve the needs of the United States as civil servants and volunteers at
election sites. Indeed, this study can be seen as a test of a particular
assumption that veterans have about themselves---the assumption that
their mere presence and civic engagement will inspire confidence in
elections among the public.

Veterans are, after all, civilians. Civilians with shared (but varied)
experience and status of prior military service, but who are no more or
less qualified for this civil service\footnote{Research is limited and
  mixed on whether veterans in general are more inclined, however
  (\citeproc{ref-nesbit2011}{Nesbit and Reingold 2011}).} over
non-veteran civilians. To be frank, announcements about concerted
efforts to recruit veterans as election workers should not reasonably
influence one's trust that votes will be counted more accurately, or
that election fraud will not occur; there is no known evidence
demonstrating that actually employing veterans as election staff or
volunteers further improves the safety, security, and overall integrity
of elections in the United States. Moreover, there's no known evidence
of positive influence on public perceptions of said outcomes when voters
encounter military veterans working as staff or volunteers at elections
sites. So, why veterans?

Two considerations underpin the notion that information about veterans
will improve trust and quell distrust in elections. First, positive
public regard for veterans \emph{in general} is sufficient to elicit
greater trust and lower distrust in elections despite the fact that
information about veterans as election workers, or even merely
information on efforts to recruit veterans, has no bearing on the actual
integrity of the electoral process.

An underlying assumption here is that there \emph{is} something special
about veterans in the eyes of the public generally, but this doesn't
include a proposition on exactly \emph{what} that special something
might be in theory. This is partly due to there being limited research
on public opinion about former and active military service members
distinguished separately from public opinion about the military as an
institution in general. Support for the military is generally pretty
high, but attitudes about the military in general can't be construed as
public opinion about veterans as a group. The evidence that can be
gathered concerning public regard for military veterans shows that,
generally, veterans are perceived favorably among the general public and
all groups express exaggerated support for veterans in more than one way
(\citeproc{ref-kleykamp2018}{Kleykamp, Hipes, and MacLean 2018}),
suggesting that public regard for veterans is prescribed by widespread
social norm. Although military recruitment shows a downward trend as of
late, public perceptions of veterans are overwhelmingly positive
(\citeproc{ref-kleykamp2023}{Kleykamp, Schwam, and Wenig 2023}).

Beyond such findings, it is overtly clear that former and active
military service members are highly revered nationwide. Veteran's day is
a national holiday, and custom prescribes thanking military members for
their service. Calls to ``support the troops'' are distinct from calls
to support the politics the troops are deployed to pursue. In 2016,
former NFL player Colin Kaepernick went from sitting to kneeling in
protest during the playing of the National Anthem upon criticism that
sitting in protest implied disrespect for active and former military
service members (\citeproc{ref-haerens2018}{Haerens 2018}).

Generally high public regard for veterans is taken together with
literature suggesting that public trust in elections is malleable to
factors that range in logical relevance to the actual structure or
administration of elections. On the one hand, public trust is, indeed,
influenced by information specifically about the processes of conducting
elections (\citeproc{ref-gaudette2025}{Gaudette et al. 2025}). On the
other hand, the public's trust in elections is not beholden to the
demonstrated \emph{trustworthiness} of the institution of election
administration. Whether generally high public reverence for veterans is
due to injunctive social norms or not, what may be posited is simply
that positive regard for veterans in general may function sufficiently
as a cue to elicit greater trust in elections despite relevance.

Second, veterans are a particularly potent group where mention of one's
veteran status seems to have a calming effect on attitudes. More so than
just the general positive favor afforded to veterans, the group appears
to have a somewhat pacifying effect on negative or otherwise
contemptible discrimination, prejudice, and social stigma. For instance,
recent research shows that, during his campaign in the 2020 Democratic
Primaries, Pete Buttigieg's military background mitigated discrimination
against him when he was presented as a veteran married to a man
(\citeproc{ref-magni2024}{Magni and Reynolds 2024}). Similar research
has also found that a candidate's veteran status affords them better
evaluations regarding competency in particular issue areas (e.g., war
competence) (\citeproc{ref-hardy2019}{Hardy et al. 2019};
\citeproc{ref-teigen2013}{Teigen 2013}). Moreover, veteran status seems
to mitigate or nullify usual stigmas associated with mental illness.
That is to say, there is negative stigma associated with mental illness
(\citeproc{ref-corrigan2002}{Corrigan et al. 2002}) and such stigma
incurs labor market discrimination (\citeproc{ref-hipes2016}{Hipes et
al. 2016}), but evidence suggests that veteran status overrides such
stigma and discrimination (\citeproc{ref-maclean2014}{MacLean and
Kleykamp 2014}). Or, in another light, mental illness is seemingly more
\emph{understandable} (i.e., permissible) for veterans given the
presumptive reasons for their mental strife. And media framing as such
plays an important role on public perception
(\citeproc{ref-kleykamp2015}{Kleykamp and Hipes 2015}). Relatedly,
attitudes about homelessness and PTSD---ranging from perceived causes of
homelessness, whether the federal government should dedicate more
resources, perceived effectiveness of policies, to compassion for
homeless individuals---are generally more favorable for the homeless
population among veterans than for the homeless population of general
adults (\citeproc{ref-tsai2021}{Tsai et al. 2021}).

Such effects primarily pertain to attitudes about individual people or
groups (e.g., a political candidate, prospective employees, the homeless
population). The capacity of one's veteran status to pacify prejudice,
discrimination, and stigma likely extends out to veterans as a group in
a similar fashion. Information that election sites will be staffed by
veterans may be enough to ease insecurities about the competence and
commitment of the institution to conduct fair and accurate elections.

@@

This theory is constrained to public confidence in the administration of
elections in the United States. To be clear, the scope conditions in
which this theory applies concern public trust and distrust in election
administration in the anticipatory period prior to the election event.
The theoretical relationship between trust and distrust is important to
understand, especially when it comes to the dependent outcome of
interest. Before going further, it is necessary to explicate the
conception \emph{confidence in elections} with respect to how I've
defined trust and distrust in elections.

Trust is, after all, merely \emph{belief} held in the face of
uncertainty. Trust is reliance on the veracity of belief in the face of
uncertainty. Said another way, trust is \emph{assured belief} dependent
on its supposed veracity; a state of certainty with respect to some
belief or judgement. It is ``\emph{assured}'' because it is one's belief
that relies on something rather than nothing. Hence quantitative methods
of the sciences are replete with things such as \emph{confidence}
intervals around our estimates. Therefore, I understand trust to refer
to the state or quality of certainty around one's beliefs or
expectations; the extent to which one is assured of their expectations
regarding what is or what will be the case.

@@
